---
title: "Root Growth GAM analysis - First steps"

author: Veta Bonnewell

date: "2/10/2022"
output: 
  word_document:
    reference_docx: Styles_Template.docx
---
<!---  Commenting out default R Markdown Header
output:
  word_document: default
  html_document: default
  
To change default formatting
Edit Styles_Template.docx
I modified Header1-3 to reduce font size and spacing.
Not clear how to reduce font size of output and maintain formating

Use following to create simple header
--->

Title: "Root Growth GAM analysis - First steps"
Author: Veta Bonnewell
Date: "2/10/2022"


<!---  Comment  old header and  out knit instructions
###Script for Analysis of Root Growth 2019- 2020

###Goal explore GAM for modeling data

###Start:December 28, 2021

###Veta Bonnewell

#Rmarkdown Instructions
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

VB: set comment = NA (default is '##')
prompt = '>'
fig.height and fig.width in inches
Set these as chunk defaults?
Example:opts_chunk$set(message=FALSE)
set within a chunk
Unsuccessful in setting hxw = 2x3
--->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA, prompt = '>' )
```


## Read Raw Data and create data summaries for plot, tube by year, entire interval


```{r , echo=FALSE, collapse=TRUE, message = FALSE, warning=FALSE}

source("Scripts/Root-Raw-Data-Analysis.R", echo = FALSE)

```


Completed Read Raw Data and create data summaries for plot, tube by year, entire interval
 

## Run GAM analysis on data by tubes and by date (not normalized)



```{r, echo=FALSE, collapse=TRUE}
source("Scripts/GamAnalysis-Part1_BasicTrials.R", echo = TRUE)
```


## Run GAM analysis on normalized data for tubes and date.


Default parameters for gam():
          family = gaussian,  type of conditional distribution
          bs =tp   basis (type of smoother) is low-rank thin plate spline:
          Instead of giving a smoothing parameter value, sp=
             Set method to REML, ie.  Smoothing via restricted maximum likelihood
             

             
```{r, NormModel_Tube,echo=FALSE, collapse=TRUE}  
 tubeNorm_model_default_k <- gam ( rootCountTube_Norm ~ s(Day),  #ksets number of basis functions, i.e nodes
                        data = New_RootsByTube_NoObs1,
                        # family = 'gaussian' #default, type of conditional distribution
                        # bs ='tp'            #default, basis (type of smoother) is low-rank thin plate spline
                        method = "REML")  #Smoothing via restricted maximum likelihood, 
                                          # instead of giving a smoothing parameter value, "sp="
print(coef(tubeNorm_model_default_k))  # shows value of k was 10
summary(tubeNorm_model_default_k)
plot(tubeNorm_model_default_k,main = "GAM, method = REML, k=10, sp = 0.2235", residuals = TRUE, pch=1,  se=TRUE)

#Check smoothing parameter (s)  also g=called lambda?
tubeNorm_model_default_k$sp

gam.check(tubeNorm_model_default_k)

```
TENTATIVE: Model from Normalized Data by date (below) is the best model
  Deviance explained = 79.2%, 
  gam.check() 
     Deviance vs quantiles best fit
     Resids  vs linear pred:
     Histogram of residuals closer to Gaussian
     Response vs Fitted values:
     
```{r, NormModel_Date,echo=FALSE, collapse=TRUE}  
dateNorm_model_default_k <- gam ( rootCountDate_Norm ~ s(Day),  #ksets number of basis functions, i.e nodes
                        data = New_RootsByDate,
                        # family = 'gaussian' #default, type of conditional distribution
                        # bs ='tp'            #default, basis (type of smoother) is low-rank thin plate spline
                        method = "REML")  #Smoothing via restricted maximum likelihood, 
                                          # instead of giving a smoothing parameter value, "sp="
print(coef(dateNorm_model_default_k))  # shows value of k was 10
summary(dateNorm_model_default_k)
plot(dateNorm_model_default_k,main = "GAM, method = REML, k=10, sp = 0.0.0256", residuals = TRUE, pch=1,  se=TRUE)

#Check smoothing parameter (s)  also g=called lambda?
dateNorm_model_default_k$sp

gam.check(dateNorm_model_default_k)

 ```